---
title: "Week 2: Class Notes"
format: html
embed-resources: true
---

```{r}
#| label: "setup"
#| message: false
#| echo: false

library(tidyverse)
library(RColorBrewer)
```

## Data

Car data:

```{r}
data(mpg)
```

Texas housing data:
```{r}
data(txhousing)

# taking a subset of cities 
# don't worry about this code - we haven't learned how to do this yet!
sm_tx <- txhousing |>
  filter(city %in% c("Dallas","Fort Worth", "Austin",
                     "Houston", "El Paso"))
```


```{r}
?read_csv
?read.csv
```


## Tuesday - `ggplot2`


```{r}
ggplot(data = mpg)
```

```{r}
ggplot(data = mpg, 
       aes(x = class, y = hwy))
```

```{r}
ggplot(data = mpg, 
       aes(x = class, y = hwy)) +
  geom_jitter() +
  geom_boxplot()
```


This is an example of ______ aesthetics:
```{r}
ggplot(data = mpg, 
       mapping = aes(x = class, 
                     y = hwy)) +
  geom_boxplot()
```


This is an example of ______ aesthetics:
```{r}
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = class, 
                             y = hwy))
```


This is an example of ______ aesthetics:
```{r}
ggplot(data = mpg) +
  geom_jitter(mapping = aes(x = class, 
                             y = hwy,
                             color = class))
```


This is an example of ______ aesthetics:
```{r}
ggplot(data = mpg) +
  geom_jitter(mapping = aes(x = class, 
                             y = hwy),
               color = "steelblue")

```

## geoms
    
```{r}
ggplot(data = mpg,
       aes(x = cty,
           y = hwy,
           color = class)) +
  geom_point() +
  labs(x = "City (mpg)", y = "Highway (mpg)") +
  theme(axis.title = element_text(size = 14),
        legend.title = element_blank(),
        legend.text = element_text(size = 14))
```
                            

## Game Plans

```{r}
ggplot(data = sm_tx,
       aes(x = date, y = median, color = city)) + 
  geom_point() + 
  geom_smooth(method = "loess") + 
  labs(x = "Date",
       y = "Median Home Price",
       title = "Texas Housing Prices")
```

## Facets

```{r}
ggplot(data = sm_tx,
       aes(x = date, y = median)) + 
  geom_point() + 
  facet_grid(cols = vars(city)) +
  geom_smooth(method = "loess") + 
  labs(x = "Date",
       y = "Median Home Price",
       title = "Texas Housing Prices")
```

```{r}

ggplot(data = sm_tx,
       aes(x = date, y = median)) + 
  geom_point() + 
  facet_wrap(vars(city)) +
  geom_smooth(method = "loess") + 
  labs(x = "Date",
       y = "Median Home Price",
       title = "Texas Housing Prices")
```

##  `stat`


```{r}
ggplot(data = mpg,
       mapping = aes(x = class)) +
  geom_bar()
```

```{r}
ggplot(data = mpg,
       mapping = aes(x = class)) +
  stat_count(geom = "bar")
```

```{r}
ggplot(data = mpg,
       mapping = aes(x = class,
                     y = hwy)) +
  stat_summary(geom = "bar",
               fun = "mean") +
  scale_y_continuous(limits = c(0,45))
```

```{r}
ggplot(data = mpg,
       mapping = aes(x = class,
                     y = hwy)) +
  stat_summary(geom = "bar",
               fun = "max") +
  scale_y_continuous(limits = c(0,45))
```


## position

```{r}
ggplot(mpg, aes(x = fl, fill = drv)) + 
  geom_bar(position = "dodge")
```
```{r}
ggplot(mpg, aes(x = fl, fill = drv)) + 
  geom_bar(position = "fill")
```

```{r}
ggplot(mpg, aes(x = fl, fill = drv)) + 
  geom_bar(position = "stack")
```

## customization

```{r}
ggplot(mpg, aes(x = fl, fill = drv)) + 
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("#d73027", "#fdae61", "#4575b4"),
                    labels = c("4-wheel","front","rear")) +
  labs(x = "Fuel Type", 
       y = "Proportion", 
       fill = "Drive Train",
       title = "Fuel Type by Car Drive Train")
```

## Try it - Diamonds

```{r}
data(diamonds)
summary(diamonds)
```


## Try it - mpg

```{r}
ggplot(data = mpg,
       mapping = aes(x = cty,
                     y = hwy,
                     color = drv)) +
  geom_point() +
  facet_grid(cols = vars(year))
```



## Thursday - More Graphics


## Color

We loaded `RColorBrewer` at the beginning of this document.


You can look at pallete options using `display.brewer.all()` or by looking at
[https://colorbrewer2.org/](https://colorbrewer2.org/):

```{r}
display.brewer.all(n = 5, type = "div" , colorblindFriendly = TRUE)
```


```{r}
mycols <- brewer.pal(n = 4, name = "BuPu")
mycols
```


## Try it out - Colors

```{r}
ggplot(data = mpg) + 
  geom_jitter(mapping = aes(x = displ, y = hwy, color = cyl)) + 
  labs(x = "Engine Displacement (liters)", 
       y = "Highway MPG", 
       color = "Number of \nCylinders",
       title = "Car Efficiency")
```